---
title: "Riley_HW4"
format: html
---

Assignment Information\
ST 558\
Homework 4\
Individual\
Shannon Riley

Task 1: Conceptual Questions
--

1. Rewrite the following nested function call using baseR’s chaining operator:
```{r}
#| echo: false
#| output: false
# Load required packages
library(tidyverse)
```

```{r}
arrange(filter(select(as_tibble(iris), starts_with("Petal"), Species), Petal.Length < 1.55), Species)
```
   A) 
```{r}
iris |>
  as_tibble() |>
  select(starts_with('Petal'), Species) |>
  filter(Petal.Length < 1.55) |>
  arrange(Species)
```
  
2. What is meant by long format data and wide format data? Which do we generally prefer for statistical analysis?
   A) Long format data means that each observation has only one data value, while wide format data means that a single observation can have multiple data values. Long formt data is generally preferred for statistical analysis.


Task 2: 
--


#### Glass Data

Read in the glass data and assign variable names:
```{r}
glass <- read_delim('https://www4.stat.ncsu.edu/~online/datasets/glass.data', 
                       delim = ',',
                       col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'glassType')
                    )
```

Manipulate the 'glass' data object by creating a new column with 'glass type' as a factor, then filtering to only retain observations where iron weight percent is less than 0.2 and glass type is either 'tableware' or 'headlamp':
```{r}
glass |>
  mutate(
    fGlassType = factor(glassType, 
                        labels = c('buildingWindowFloat', 'buildingWindowNonFloat', 'vehicleWindowFloat', 'container', 'tableware', 'headlamps'))
  ) |>
  filter(Fe < 0.2, fGlassType == 'tableware' | fGlassType == 'headlamp')
```


##### Yeast Data

Read in the yeast data and assign variable names:
```{r}
yeast <- read_table('https://www4.stat.ncsu.edu/~online/datasets/yeast.data',
                    col_names = c('seq_name', 'mcg', 'gvh', 'alm', 'mit', 'erl', 'pox', 'vac', 'nuc', 'class')
                    )
```

Manipulate the 'yeast' data object by removing the variables 'seq_name' and 'nuc', then add columns corresponding to the mean and median of each numeric variable at each class grouping:
```{r}
yeast |>
  select(!c(seq_name, nuc)) |>
  group_by(class) |>
  mutate(
    across(
      where(is.numeric),
      list(mean = mean, median = median),
      .names = '{.col}.{fn}'
      )
    )
```


#### Wine Data

```{r}
#| echo: false
#| output: false
# Load required packages
library(readxl)
```

Import the data from the first sheet of the white wine data set:
```{r}
whiteWine <- read_excel('../Data/HW3/white-wine.xlsx', sheet = 1)
whiteWine
```

Read in the variable names from the second sheet of the white wine Excel file:
```{r}
whiteWineNames <- read_excel('../Data/HW3/white-wine.xlsx', sheet = 2)
```

Use this tibble of variable names to overwrite the column names in the 'whiteWine' tibble and add a column to this data set to indicate that the wines are white:
```{r}
colnames(whiteWine) <- c(whiteWineNames$Variables)
whiteWine$type <- 'white'
```

Read in the red wine data set, replace variable names using the same set of variable names as above, and add a column to this data set to indicate that the wines are red:
```{r}
redWine <- read_delim('https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv', delim = ';')
colnames(redWine) <- c(whiteWineNames$Variables)
redWine$type <- 'red'
```

Combine the two data sets:
```{r}
wine <- bind_rows(whiteWine, redWine)
```

Manipulate the 'wine' data object by filtering for high-quality wines that have a reasonable alcohol value, selecting a subset of the variables, and creating mean and standard deviation columns of the 'alcohol' variable for each value of the 'quality' variable:
```{r}
wine |>
  filter(quality > 6.5, alcohol < 13.2) |>
  arrange(desc(quality)) |>
  select(contains('acid'), alcohol, type, quality) |>
  group_by(quality) |>
  mutate(
    across(alcohol,
      list(mean = mean, sd = sd),
      .names = '{.col}.{fn}'
      )
    )
```

Task 3:
--

```{r}
#| echo: false
#| output: false
# Load required packages
library(RSQLite)
library(DBI)
```

Connect to the 'Lahman' database and look at all of the tables in the database:
```{r}
con <- dbConnect(RSQLite::SQLite(), '../Data/HW4/lahman.db')
dbListTables(con)
```

Use the 'tbl()' function and 'dplyr' package to return of the data from the 'Teams' table for the year 2015:
```{r}
tbl(con, 'Teams') |>
  filter(yearID == 2015) |>
  show_query() |>
  collect()
```

Use the 'tbl()' function and the 'sql()' function within that function to return of the data from the 'Teams' table for the year 2015:
```{r}
tbl(con, sql(
'SELECT *
FROM `Teams`
WHERE (`yearID` = 2015.0)')
) |>
  collect()
```

Return all of the players in the hall of fame, the year they were voted into the hall of fame, and their
category:
```{r}
tbl(con, 'HallOfFame') |>
  filter(inducted == 'Y') |>
  select(c(playerID, yearID, category)) |>
  collect()
```

Combine the previous table with the 'People' table in order to have the first and last names added to what was returned above:
```{r}
hallPlayers <- inner_join(tbl(con, 'HallOfFame') |>
                            filter(inducted == 'Y') |>
                            select(c(playerID, yearID, category)),
                          tbl(con, 'People') |>
                            select(c(playerID, nameFirst, nameLast)),
                          by = join_by(playerID == playerID)
                          ) |>
  collect()
hallPlayers
```

Return the 'playerID' (managerID), 'G', 'W', and 'L' columns from the 'Managers' table, then determine the overall win/loss record, then create a new variable that is the career win/loss percentage, and finally sort the resulting data by the win/loss percentage (descending):
```{r}
mngrs <- tbl(con, 'Managers') |>
  select(c(playerID, G, W, L)) |>
  group_by(playerID) |>
  summarize(G_managed = sum(G, na.rm = TRUE),
            Total_W = sum(W, na.rm = TRUE),
            Total_L = sum(L, na.rm = TRUE)) |>
  collect() |>
  mutate(wlPercent = Total_W / G_managed) |>
  arrange(desc(wlPercent))
mngrs
```

Return information for only the people that managed a team and were inducted into the hall of fame using the results from the previous sections:
```{r}
inner_join(hallPlayers, mngrs,
           by = join_by(playerID == playerID))

```

Disconnect from the database:
```{r}
dbDisconnect(con)
```

