---
title: "Homework 2: Group"
author: "Group 4, Combined by MMendoza"
format: html
editor: visual
---

# Cody Ashby (A)

## Task One: Basic Vector Practice

Below is some hypothetical data that measures the systolic blood pressure of 20 subjects, both before and after a treatment. Labels has been included as well as the difference between both measurements for each subject.

```{r}
pre_bp_1 = c(130,128,116,124,133,134,118,126,114,127,141,138,128,140,137,131,120,128,139,135)
post_bp_1 = c(114,98,113,97,107,116,113,111,119,117,101,119,130,122,106,106,124,102,117,113)
patients_1 = paste("Subject",1:20,sep="_")
diff_bp_1 = pre_bp_1 - post_bp_1
```

Here are the mean of the differences:

```{r}
mean(diff_bp_1)
```

The subjects that have a drop in blood pressure after the treatment are shown below:

```{r}
which(diff_bp_1 > 0)
```

The values associated with these subjects are as follows:

```{r}
diff_bp_1[diff_bp_1 > 0]
```

Also included are the average of these positive differences:

```{r}
mean(diff_bp_1[diff_bp_1 > 0])
```

## Task Two: Basic Data Frame Practice

We can combine these data vectors into a data frame, as shown below:

```{r}
bp_data_1=data.frame(patients_1,pre_bp_1,post_bp_1,diff_bp_1)
```

Let's see whose blood pressure actually *increased* after the treatment.

```{r}
subset(bp_data_1,diff_bp_1<0)
```

To determine which patients ended up with a normal blood pressure measurement after the treatment, we'll create a new column:

```{r}
normal_bp_1=bp_data_1[,3] < 120
```

With this new column, we can attach it to our previous data frame to get an "updated" version of it.

```{r}
updated_bp_data_1=data.frame(bp_data_1,normal_bp_1)
knitr::kable(updated_bp_data_1)
```

## Task Three: List Practice

We follow a similar process to create a data frame for a second set of subjects that have received a placebo instead.

```{r}
pre_bp_2=c(138,135,147,117,152,134,114,121,131,130)
post_bp_2=c(105,136,123,130,134,143,135,139,120,124)
diff_bp_2=pre_bp_2 - post_bp_2
patients_2=paste("Subject",21:30,sep="_")
bp_data_2=data.frame(patients_2,pre_bp_2,post_bp_2,diff_bp_2)
normal_bp_2=bp_data_2[,3] < 120
updated_bp_data_2=data.frame(bp_data_2,normal_bp_2)
```

The two data frames will now be combined using a list technique:

```{r}
overall_bp_data=list(treatment=updated_bp_data_1,placebo=updated_bp_data_2)
```

Below are three different ways to access the first element of this list (those that received the treatment):

```{r}
# using the dollar sign
overall_bp_data$treatment
```

```{r}
# using the single bracket
overall_bp_data[1]
```

```{r}
# using the double bracket
overall_bp_data[[1]]
```

We can access the initial measurements from the placebo data by using the double bracket to specify the data set and the `[i,j]` notation for the second column of it:

```{r}
overall_bp_data[[2]][,2]
```

## Task Four: Control Flow Practice

Now is the opportunity to utilize conditional logic; before diving in, let's define a new character column for each data frame

```{r}
updated_bp_data_1$status <- character(20)
updated_bp_data_2$status <- character(10)
```

We'll create a `for` loop while using `if/else` logic to classify the type of blood pressure measurements after receiving the treatment in the first group:

```{r}
for (i in 1:20){
  if (updated_bp_data_1$post_bp_1[i] <= 120){
    updated_bp_data_1$status[i] = 'optimal'
  } else if ((updated_bp_data_1$post_bp_1[i] > 120) & (updated_bp_data_1$post_bp_1[i] <= 130)){
    updated_bp_data_1$status[i] = 'borderline'
  } else if (updated_bp_data_1$post_bp_1[i] > 130){
    updated_bp_data_1$status[i] = 'high'
  }
}
```

The same process will be used for those in the placebo group:

```{r}
for (i in 1:10){
  if (updated_bp_data_2$post_bp_2[i] <= 120){
    updated_bp_data_2$status[i] = 'optimal'
  } else if ((updated_bp_data_2$post_bp_2[i] > 120) & (updated_bp_data_2$post_bp_2[i] <= 130)){
    updated_bp_data_2$status[i] = 'borderline'
  } else if (updated_bp_data_2$post_bp_2[i] > 130){
    updated_bp_data_2$status[i] = 'high'
  }
}
```

Here's the final data frame:

```{r}
overall_bp_data = list(updated_bp_data_1,updated_bp_data_2)
overall_bp_data
```

# **Michael Mendoza (B)**

# Question 1

The following code creates the pre-/post-treatment records as 2 different vectors

```{r}

pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
print(pre_bp)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
print(post_bp)

```

The following Code assigns names to these vectors and creates a vector of these names.

```{r}
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
print(pre_bp)

names(post_bp) <- paste("Subject", 1:20, sep = "_")
print(post_bp)

patient <- paste("Subject", 1:20, sep = "_")
print(patient)
```

The following stores the difference of the pre-/post- treatment records, a positive record indicates a decrease in blood pressure.

```{r}
diff_bp <- pre_bp - post_bp
print(diff_bp)
```

This is the mean of the difference of among all patients.

```{r}

print(mean(diff_bp))
```

The following returns a vector of which subjects experienced a positive change, and stores the records as a subset.

```{r}

print(which(diff_bp>0))

pos_subset <- diff_bp[which(diff_bp>0)]
```

This is the mean of the difference of among patients who experienced a decrease in blood pressure.

```{r}
print(mean(diff_bp))
```

# Question 2

The following code creates a data frame with 4 variables: patient, pre_bp, post_bp, and diff_bp

```{r}
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
```

Return only negative difference patients.

```{r}
print(bp_df[bp_df$diff_bp < 0, ])
```

Adding True/False flag to dataset when post_bp \< 120

```{r}
bp_df$lt_120 <- bp_df$diff_bp < 0
print(bp_df)
```

Knit code to print bp_df to final document.

```{r}
knitr::kable(bp_df, caption = 'Blood Pressure Data Table')
```

# Question 3

Create the new dataset.

```{r}
pre_bp2 <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

post_bp2 <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

names(pre_bp2) <- paste("Subject", 1:10, sep = "_")

names(post_bp2) <- paste("Subject", 1:10, sep = "_")

patient2 <- paste("Subject", 1:10, sep = "_")

diff_bp2 <- pre_bp2 - post_bp2

bp_df2 <- data.frame("patient" = patient2, "pre_bp" = pre_bp2, "post_bp" = post_bp2, "diff_bp" = diff_bp2)

bp_df2$lt_120 <- bp_df2$diff_bp < 0
```

Create the list with 2 elements: treatment, placebo.

```{r}
bp_list <- list("treatment" = bp_df, "placebo" = bp_df2)
```

3 different types of syntax to access the first list element.

```{r}

bp_list[[1]]

bp_list["treatment"]

head(bp_list, 1)

```

Access Placebo: pre_bp in a single line.

```{r}
print(bp_list$placebo$pre_bp)
```

# Question 4

Create new column named status in each dataset in listing

```{r}
bp_list$treatment$status <- character(20)

bp_list$placebo$status <- character(10)
```

Using if/else lolgic to derive status flag in treatment dataframe.

```{r}
for (i in 1:20) {
  
  if (bp_list$treatment$post_bp[i] <= 120) {
    bp_list$treatment$status[i] <- "optimal"
  }
    else if (bp_list$treatment$post_bp[i] > 130) {
      bp_list$treatment$status[i] <- "high"
    }
    else {
      bp_list$treatment$status[i] <- "borderline"
    }
}

print(bp_list$treatment)
```

Repeat for placebo data

```{r}
for (i in 1:10) {
  
  if (bp_list$placebo$post_bp[i] <= 120) {
    bp_list$placebo$status[i] <- "optimal"
  }
  else if (bp_list$placebo$post_bp[i] > 130) {
    bp_list$placebo$status[i] <- "high"
  }
  else {
    bp_list$placebo$status[i] <- "borderline"
  }
}

print(bp_list$placebo)
```

# **Shannon Riley (C)**

# Tasks

## \## Task 1: Basic Vector Practice

Create vectors of pre-treatment and post-treament blood pressure measurements:

```{r}
preBP <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

postBP <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

Assign subject names to the blood pressure measurements (vector elements):

```{r}
names(preBP) <- paste('Subject', 1:20, sep = "_")
names(postBP) <- paste('Subject', 1:20, sep = "_")
```

Calculate the change in blood pressure for each patient:

```{r}
changeBP <- preBP - postBP
```

Calculate average decrease in blood pressure across all patients:

```{r}
avgChange <- mean(changeBP)
```

Determine which patients experienced a decrease in blood pressure:

```{r}
dcrs <- which(changeBP > 0)
```

Subset the vector of differences to only retain those that had a decrease in blood pressure:

```{r}
onlyDecr <- changeBP[dcrs]
```

Calculate the average decrease in blood pressure for patients whose blood pressure decreased:

```{r}
avgDecr <- mean(onlyDecr)
```

## \## Task 2: Basic Data Frame Practice

Create a data frame with columns for patient, pre-treamtent blood pressure, post-treamtent blood pressure, and difference in blood pressure:

```{r}
bp <- data.frame(patient = names(preBP), preBP, postBP, changeBP, row.names = NULL)
```

Subset the data frame to retain only observations where the difference in blood pressure is negative:

```{r}
subset(bp, changeBP < 0)
```

Add a new column to the data frame with TRUE is post-treatment blood pressure is less than 120:

```{r}
bp$postBP120 <- c(bp$postBP < 120)
```

```{r}
#| echo: false
knitr::kable(bp, col.names = c('Patient Identifier', 'Pre-Treatment Blood Pressure', 'Post-Treatment Blood Pressure', 'Change in Blood Pressure', 'Post-Treamtent Blood Pressure Less Than 120?'), align = c('l', 'r', 'r', 'r', 'r'), caption = 'Blood Pressure Change')
```

## \## Task 3: List Practice

Create a data frame of blood pressure measurements before and after a placebo treatment, the difference between measurements, and whether or not the post-treatment measurement was below 120:

```{r}
plcbPreBP <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)
plcbPostBP <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)
placebo <- data.frame(patient = paste('Subject', 1:10, sep = "_"), preBP = plcbPreBP, postBP = plcbPostBP, changeBP = plcbPreBP - plcbPostBP, postBP120 = c(plcbPostBP < 120))
```

Create and store a list with both data frames:

```{r}
bpList <- list(treatment = bp, placebo = placebo)
```

Access the first list element using three different types of syntax:

```{r}
bpList$treatment
bpList[1]
bpList[[1]]
```

Access the pre-treatment measurement column of the placebo data frame in one line:

```{r}
bpList$placebo[2]
```

## \## Task 4: Control Flow Practice

Create a new column in each data frame called status:

```{r}
bpList$treatment$status <- character(20)
bpList$placebo$status <- character(10)
```

Use a for loop and if/else logic to fill the status column of the treatment data frame with "optimal", "borderline", or "high" based on the post-treatment blood pressure measurement:

```{r}
for (patient in 1:dim(bpList$treatment)[1]) {
  if (bpList$treatment$postBP[patient] <= 120) {
    bpList$treatment$status[patient] <- 'optimal'
    }
  else if (bpList$treatment$postBP[patient] > 130) {
    bpList$treatment$status[patient] <- 'high'
    }
  else {
    bpList$treatment$status[patient] <- 'borderline'
    }
}
```

Use a for loop and if/else logic to fill the status column of the placebo data frame with "optimal", "borderline", or "high" based on the post-treatment blood pressure measurement:

```{r}
for (patient in 1:dim(bpList$placebo)[1]) {
  if (bpList$placebo$postBP[patient] <= 120) {
    bpList$placebo$status[patient] <- 'optimal'
    }
  else if (bpList$placebo$postBP[patient] > 130) {
    bpList$placebo$status[patient] <- 'high'
    }
  else {
    bpList$placebo$status[patient] <- 'borderline'
    }
}
```

# Stanley Xiao (D)

## Task 1: Basic Vector Practice

1.  Assigning names for pre-treatment BP values, post-treatment BP values

```{r}
#Pre-Treatment BP values
pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

#Post-Treatment BP values
post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

2.  Assigning Subject Names for each Pre/Post-treatment BP values

```{r}
names(pre_bp) <- paste("Subject", 1:20, sep="_")
names(post_bp) <- paste("Subject", 1:20, sep="_")
```

3.  Change in BP for each Patient

```{r}
diff_bp <- pre_bp - post_bp
```

4.  Average Decrease in BP Across All Patients

```{r}
avg_change <- mean(diff_bp)
```

5.  Which Patients Experienced Decrease in BP After Treatment (Positive Change)

```{r}
lower_bp <- which(diff_bp > 0)
```

6.  Subset Vector in \[HW2 1.5\] to Only Return Positive Change

```{r}
only_lower <- diff_bp[lower_bp]
```

7.  Average Decrease in BP for Positive Change (BP Decrease)

```{r}
avg_lower <- mean(only_lower)
```

## Task 2: Basic Data Frame Practice

1.  Create a data frame object with four columns corresponding to your data above: patient, pre_bp, post_bp, and diff_bp

```{r}
bp <- data.frame(patient = names(pre_bp), pre_bp, post_bp, diff_bp, row.names=NULL)

bp
```

2.  Return only rows where the diff_bp column is negative. (Use \[ or learn about the subset() function if you’d like. If you use \[, don’t reference the original vector from the first part, access the column of the data frame when looking at making a comparison.

```{r}
subset(bp, diff_bp < 0)
```

3.  Add a new column to the data frame corresponding to TRUE if the post_bp is less than 120. Recall you can use \$ to reference a column. If you reference a column that doesn’t exist, and save a vector (of appropriate length in it), that vector becomes a column of your data frame! Similar to the previous question, don’t reference the original vector from the first part, access the column of the data frame when looking at making a comparison.

```{r}
bp$bp_post_lt_120 <- bp$post_bp < 120
```

4.  Finally, print the data frame out nicely in your final document by modifying the code below appropriately.

```{r}
knitr::kable(bp)
```

## Task 3: List Practice

1.  Create a new data frame with this data in it that is similar to the data frame from task 2 (including the new column).

```{r}
#Now we're adding in another experiment where the 'treatment' was actually a placebo.
placebo_bp_pre <- c(138, 135, 147, 117, 152, 134, 114, 121, 131, 130)

placebo_bp_post <- c(105, 136, 123, 130, 134, 143, 135, 139, 120, 124)

placebo <- data.frame(patient=paste('Subject', 1:10, sep="_"), pre_bp = placebo_bp_pre, post_bp = placebo_bp_post, diff_bp = placebo_bp_pre - placebo_bp_post, bp_post_lt_120 = placebo_bp_post < 120)
```

2.  Now create and store a list with two elements: • 1st element named treatment and contains the first data frame you created. • 2nd element named placebo and contains the second data frame you created.

```{r}
bp_list <- list(treatment = bp, placebo = placebo)
```

3.  Access the first list element using three different types of syntax.

```{r}
# Method 1
bp_list$treatment
# Method 2
bp_list[["treatment"]]
# Method 3
bp_list[[1]]
```

4.  In one line, access the placebo data frame, pre_bp column.

```{r}
bp_list$placebo$pre_bp
```

## Task 4: Control Flow Practice

1.  Suppose we want to characterize the post-treatment (or placebo) blood pressure measurement as optimal (≤ 120), borderline (120 \< bp ≤ 130), and high (\> 130). First, create a new column in each data frame from above called status. You can do this via
    1.  your_df\$status \<- character(20) #or 10 depending on number of observations

```{r}
bp_list$treatment$status <- character(20)
bp_list$placebo$status <- character(10)
```

2.  For the non-placebo data frame (within the list), create a for loop and use if/then/else logic to create the status column’s values.

```{r}
for (i in 1:nrow(bp_list$treatment)) {
  if (bp_list$treatment$bp_post[i] <= 120) {
    bp_list$treatment$status[i] <- "Optimal"
  } else if (bp_list$treatment$bp_post[i] <= 130) {
    bp_list$treatment$status[i] <- "Borderline"
  } else {
    bp_list$treatment$status[i] <- "High"
  }
}
```

3.  Repeat for the placebo data frame (within the list).

```{r}
for (i in 1:nrow(bp_list$placebo)) {
  if (bp_list$placebo$bp_post[i] <=  120) {
    bp_list$placebo$status[i] <- "Optimal"
  } else if (bp_list$placebo$bp_post[i] <= 130) {
    bp_list$placebo$status[i] <- "Borderline"
  } else {
    bp_list$placebo$status[i] <- "High"
  }
}
```
