---
title: "Homework 2: Individual"
author: "Michael Mendoza"
format: html
editor: visual
date: 31AUG2026
purpose: create HW2 content (individual)
---

# Question 1

The following code creates the pre-/post-treatment records as 2 different vectors

```{r}

pre_bp <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
print(pre_bp)

post_bp <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
print(post_bp)

```

The following Code assigns names to these vectors and creates a vector of these names.

```{r}
names(pre_bp) <- paste("Subject", 1:20, sep = "_")
print(pre_bp)

names(post_bp) <- paste("Subject", 1:20, sep = "_")
print(post_bp)

patient <- paste("Subject", 1:20, sep = "_")
print(patient)
```

The following stores the difference of the pre-/post- treatment records, a positive record indicates a decrease in blood pressure.

```{r}
diff_bp <- pre_bp - post_bp
print(diff_bp)
```

This is the mean of the difference of among all patients.

```{r}

print(mean(diff_bp))
```

The following returns a vector of which subjects experienced a positive change, and stores the records as a subset.

```{r}

print(which(diff_bp>0))

pos_subset <- diff_bp[which(diff_bp>0)]
```

This is the mean of the difference of among patients who experienced a decrease in blood pressure.

```{r}
print(mean(diff_bp))
```

# Question 2

The following code creates a data frame with 4 variables: patient, pre_bp, post_bp, and diff_bp

```{r}
bp_df <- data.frame(patient, pre_bp, post_bp, diff_bp)
```

Return only negative difference patients.

```{r}
print(bp_df[bp_df$diff_bp < 0, ])
```

Adding True/False flag to dataset when post_bp \< 120

```{r}
bp_df$lt_120 <- bp_df$diff_bp < 0
print(bp_df)
```

Knit code to print bp_df to final document.

```{r}
knitr::kable(bp_df, caption = 'Blood Pressure Data Table')
```
