---
title: "Homework 2: Individual"
author: "Michael Mendoza"
format: html
editor: visual
date: 31AUG2026
purpose: create HW2 content (individual)
---

# Question 1

The following code creates the pre-/post-treatment records as 2 different vectors

print(mean(pos_subset))

```{r}

measurement_pre <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)
print(measurement_pre)

measurement_post <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
print(measurement_post)

```

The following Code assigns names to these vectors

```{r}
names(measurement_pre) <- paste("Subject", 1:20, sep = "_")
print(measurement_pre)

names(measurement_post) <- paste("Subject", 1:20, sep = "_")
print(measurement_post)
```

The following stores the difference of the pre-/post- treatment records, a positive record indicates a decrease in blood pressure.

```{r}
diffs <- measurement_pre - measurement_post
print(diffs)
```

This is the mean of the difference of among all patients.

```{r}

print(mean(diffs))
```

The following returns a vector of which subjects experienced a positive change, and stores the records as a subset.

```{r}

print(which(diffs>0))

pos_subset <- diffs[which(diffs>0)]
```

This is the mean of the difference of among patients who experienced a decrease in blood pressure.

```{r}
print(mean(pos_subset))
```
