---
title: "Riley_HW2"
author: "Shannon Riley"
format: html
---

# Assignment Information
ST 558\
Homework 2\
Individual Portion\
Shannon Riley

# Tasks

## Task 1: Basic Vector Practice
--

Create vectors of pre-treatment and post-treament blood pressure measurements:
```{r}
preBP <- c(130, 128, 116, 124, 133, 134, 118, 126, 114, 127, 141, 138, 128, 140, 137, 131, 120, 128, 139, 135)

postBP <- c(114, 98, 113, 99, 107, 116, 113, 111, 119, 117, 101, 119, 130, 122, 106, 106, 124, 102, 117, 113)
```

Assign subject names to the blood pressure measurements (vector elements):
```{r}
names(preBP) <- paste('Subject', 1:20, sep = "_")
names(postBP) <- paste('Subject', 1:20, sep = "_")
```

Calculate the change in blood pressure for each patient:
```{r}
changeBP <- preBP - postBP
```

Calculate average decrease in blood pressure across all patients:
```{r}
avgChange <- mean(changeBP)
```

Determine which patients experienced a decrease in blood pressure:
```{r}
dcrs <- which(changeBP > 0)
```

Subset the vector of differences to only retain those that had a decrease in blood pressure:
```{r}
onlyDecr <- changeBP[dcrs]
```

Calculate the average decrease in blood pressure for patients whose blood pressure decreased:
```{r}
avgDecr <- mean(onlyDecr)
```


## Task 2: Basic Data Frame Practice
--

Create a data frame with columns for patient, pre-treamtent blood pressure, post-treamtent blood pressure, and difference in blood pressure:
```{r}
bp <- data.frame(patient = names(preBP), preBP, postBP, changeBP, row.names = NULL)
```

Subset the data frame to retain only observations where the difference in blood pressure is negative:
```{r}
subset(bp, changeBP < 0)
```

Add a new column to the data frame with TRUE is post-treatment blood pressure is less than 120:
```{r}
bp$postBP120 <- c(bp$postBP < 120)
```

```{r}
#| echo: false
knitr::kable(bp, col.names = c('Patient Identifier', 'Pre-Treatment Blood Pressure', 'Post-Treatment Blood Pressure', 'Change in Blood Pressure', 'Post-Treamtent Blood Pressure Less Than 120?'), align = c('l', 'r', 'r', 'r', 'r'), caption = 'Blood Pressure Change')
```

