---
title: "Practice with Basic R Data Structures and Control Flow"
author: "Cody Ashby"
format: html
editor: visual
---

## Task One: Basic Vector Practice

Below is some hypothetical data that measures the systolic blood pressure of 20 subjects, both before and after a treatment. Labels has been included as well as the difference between both measurements for each subject.

```{r}
pre_bp_1 = c(130,128,116,124,133,134,118,126,114,127,141,138,128,140,137,131,120,128,139,135)
post_bp_1 = c(114,98,113,97,107,116,113,111,119,117,101,119,130,122,106,106,124,102,117,113)
patients_1 = paste("Subject",1:20,sep="_")
diff_bp_1 = pre_bp_1 - post_bp_1
```

Here are the mean of the differences:

```{r}
mean(diff_bp_1)
```

The subjects that have a drop in blood pressure after the treatment are shown below:

```{r}
which(diff_bp_1 > 0)
```

The values associated with these subjects are as follows:

```{r}
diff_bp_1[diff_bp_1 > 0]
```

Also included are the average of these positive differences:

```{r}
mean(diff_bp_1[diff_bp_1 > 0])
```

## Task Two: Basic Data Frame Practice

We can combine these data vectors into a data frame, as shown below:

```{r}
bp_data_1=data.frame(patients_1,pre_bp_1,post_bp_1,diff_bp_1)
```

Let's see whose blood pressure actually *increased* after the treatment.

```{r}
subset(bp_data_1,diff_bp_1<0)
```

To determine which patients ended up with a normal blood pressure measurement after the treatment, we'll create a new column:

```{r}
normal_bp_1=bp_data_1[,3] < 120
```

With this new column, we can attach it to our previous data frame to get an "updated" version of it.

```{r}
updated_bp_data_1=data.frame(bp_data_1,normal_bp_1)
knitr::kable(updated_bp_data_1)
```

## Task Three: List Practice

We follow a similar process to create a data frame for a second set of subjects that have received a placebo instead.

```{r}
pre_bp_2=c(138,135,147,117,152,134,114,121,131,130)
post_bp_2=c(105,136,123,130,134,143,135,139,120,124)
diff_bp_2=pre_bp_2 - post_bp_2
patients_2=paste("Subject",21:30,sep="_")
bp_data_2=data.frame(patients_2,pre_bp_2,post_bp_2,diff_bp_2)
normal_bp_2=bp_data_2[,3] < 120
updated_bp_data_2=data.frame(bp_data_2,normal_bp_2)
```

The two data frames will now be combined using a list technique:

```{r}
overall_bp_data=list(treatment=updated_bp_data_1,placebo=updated_bp_data_2)
```

Below are three different ways to access the first element of this list (those that received the treatment):

```{r}
# using the dollar sign
overall_bp_data$treatment
```

```{r}
# using the single bracket
overall_bp_data[1]
```

```{r}
# using the double bracket
overall_bp_data[[1]]
```

We can access the initial measurements from the placebo data by using the double bracket to specify the data set and the `[i,j]` notation for the second column of it:

```{r}
overall_bp_data[[2]][,2]
```

## Task Four: Control Flow Practice

Now is the opportunity to utilize conditional logic; before diving in, let's define a new character column for each data frame

```{r}
updated_bp_data_1$status <- character(20)
updated_bp_data_2$status <- character(10)
```

We'll create a `for` loop while using `if/else` logic to classify the type of blood pressure measurements after receiving the treatment in the first group:

```{r}
for (i in 1:20){
  if (updated_bp_data_1$post_bp_1[i] <= 120){
    updated_bp_data_1$status[i] = 'optimal'
  } else if ((updated_bp_data_1$post_bp_1[i] > 120) & (updated_bp_data_1$post_bp_1[i] <= 130)){
    updated_bp_data_1$status[i] = 'borderline'
  } else if (updated_bp_data_1$post_bp_1[i] > 130){
    updated_bp_data_1$status[i] = 'high'
  }
}
```

The same process will be used for those in the placebo group:

```{r}
for (i in 1:10){
  if (updated_bp_data_2$post_bp_2[i] <= 120){
    updated_bp_data_2$status[i] = 'optimal'
  } else if ((updated_bp_data_2$post_bp_2[i] > 120) & (updated_bp_data_2$post_bp_2[i] <= 130)){
    updated_bp_data_2$status[i] = 'borderline'
  } else if (updated_bp_data_2$post_bp_2[i] > 130){
    updated_bp_data_2$status[i] = 'high'
  }
}
```

Here's the final data frame:

```{r}
overall_bp_data = list(updated_bp_data_1,updated_bp_data_2)
overall_bp_data
```
