---
title: "Riley_HW3"
format: html
---

Assignment Information\
ST 558\
Homework 3\
Individual\
Shannon Riley

Task 1: Conceptual Questions
--

1. If your working directory is myfolder/homework/, what relative path would you specify to get the file
located at myfolder/MyData.csv?
   A) You need to back out of the "homework" folder because the file you want is    just in "myfolder", so you would use the relative path "../MyData.csv"
2. What are the major benefits of using R projects?
   A) The major benefit to using R projects is that it tells your code where
   to find files without needed to specify full path names, since R will look      for the file being called within the folder that the project is in. This        makes it easier to share your work. R projects also have their own workspace
   and history.
3. What is Git and what is GitHub?
   A) Git is a version control software. To use it, it needs to be installed on
   your computer. GitHub is a website that stores files. Git can work between
   GitHub and your local computer to pull files from GitHub and then push your 
   changes back to GitHub. GitHub is helpful for working with teams, but everyone needs to have Git on their computers.
4. What are the two main differences between a tibble and a data.frame?
   1) A tibble is a fancy type of data frame that has extra structure to make it easy to understand what you're looking at when you print it. The basic structure is the same as a data frame, but it prints out nicely, showing you the dimensions and automatically only shows a small subset of the rows (no need to use the *head()* function and tells you how many more rows there are.
   2) Another difference is that using *tidyverse* functions on a tibble will produce a tibble rather than simplifying down, if the output could be simplified down.

Task 2: Reading Delimited Data
--

```{r}
#| echo: false
#| output: false
# Load required packages
library(readr)
```

Read in the glass data and assign variable names
```{r}
glass <- read_delim('https://www4.stat.ncsu.edu/~online/datasets/glass.data', 
                       delim = ',',
                       col_names = c('Id', 'RI', 'Na', 'Mg', 'Al', 'Si', 'K', 'Ca', 'Ba', 'Fe', 'glassType')
)
glass
```


Read in the yeast data and assign variable names
```{r}
yeast <- read_fwf('https://www4.stat.ncsu.edu/~online/datasets/yeast.data',
                  fwf_cols('seq_name' = 12, 'mcg' = 6, 'gvh' = 6, 'alm' = 6, 'mit' = 6, 'erl' = 6, 'pox' = 6, 'vac' = 6, 'nuc' = 6, 'class' = 3)
)
yeast
```


Task 3: Combining Excel and Delimited Data
--

```{r}
#| echo: false
#| output: false
# Load required packages
library(readxl)
library(dplyr)
```
#### White Wine Data Set

Import the data from the first sheet of the white wine data set
```{r}
whiteWine <- read_excel('../Data/HW3/white-wine.xlsx', sheet = 1)
whiteWine
```


Read in the variable names from the second sheet of the white wine Excel file
```{r}
whiteWineNames <- read_excel('../Data/HW3/white-wine.xlsx', sheet = 2)
```

Use this tibble of variable names to overwrite the column names in the 'whiteWine' tibble
```{r}
colnames(whiteWine) <- c(whiteWineNames$Variables)
```

Add a column to this data set to indicate that the wines are white
```{r}
whiteWine$type <- c(rep('white', dim(whiteWine)[1]))
```

#### Red Wine Data Set

Read in the red wine data set and replace variable names using the same set of variable names as above
```{r}
redWine <- read_delim('https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv', delim = ';')
colnames(redWine) <- c(whiteWineNames$Variables)
```

Add a column to this data set to indicate that the wines are red
```{r}
redWine$type <- c(rep('red', dim(redWine)[1]))
```

#### Both Data Sets

Combine the two data sets
```{r}
wine <- bind_rows(whiteWine, redWine)
```


Task 4: Writing a Function
--

#### Interquartile Range

Write a function that takes a vector of numbers and returns the interquartile range
```{r}
getIQR <- function(inputVector) {
  if(!is.vector(inputVector) | !is.numeric(inputVector)){
    stop('Input must be a numeric vector.')
  }
  unname(quantile(inputVector)[4] - quantile(inputVector)[2])
}
```

Apply the *getIQR()* function to the 'fixed_acidity' and 'citric_acid' variables of the 'wines' data set
```{r}
getIQR(wine$fixed_acidity)
getIQR(wine$citric_acid)
```

#### Coefficient of Variation

Write a function that takes a vector of numbers and returns the sample coefficient of variation
```{r}
getCoefVar <- function(inputVector) {
  if(!is.vector(inputVector) | !is.numeric(inputVector)){
    stop('Input must be a numeric vector.')
  }
  sd(inputVector)/abs(mean(inputVector))
}
```

Apply the *getCoefVar()* function to the 'fixed_acidity' and 'citric_acid' variables of the 'wines' data set
```{r}
getCoefVar(wine$fixed_acidity)
getCoefVar(wine$citric_acid)
```

#### Summary of Both

Write a function that takes in a vector of numbers and returns both the IQR and the coefficient of variation using the functions written above
```{r}
getSummaries <- function(inputVector) {
  if(!is.vector(inputVector) | !is.numeric(inputVector)){
    stop('Input must be a numeric vector.')
  }
  iqRange <- getIQR(inputVector)
  coefVar <- getCoefVar(inputVector)
  return(list('InterquartileRange' = iqRange, 'CoefficientofVariation' = coefVar))
}
```

Apply the *getSummaries()* function to the 'fixed_acidity' and 'citric_acid' variables of the 'wines' data set
```{r}
getSummaries(wine$fixed_acidity)
getSummaries(wine$citric_acid)
```

#### Infix Function

Write an infix function that can be used to replace the *cbind()* function
```{r}
`%bind%` <- function(a, b) cbind(a, b)
```

Test the new infix function on the following data frames and then both data frames and the following vector
```{r}
# Set up data frames and vector
df1 <- data.frame(unif = runif(10), norm = rnorm(10))
df2 <- data.frame(gamma = rgamma(10, shape = 1), beta = rbeta(10, shape1 = 5, shape2 = 1))
vec1 <- rnorm(10, mean = 10)

# Test function
df1 %bind% df2
df1 %bind% df2 %bind% vec1
```
