---
title: "Homework 4"
author: "Michael Mendoza"
format: html
editor: visual
---

# Setup/Misc.

```{r}
#| echo: false

# Librarys Required
library(tidyverse)
library(readxl)
library(readr)
library(DBI)
library(RSQLite)


# Set seed for random data creation for replicability
set.seed(123)
```

# Task 1

## Q1

```{r}
as_tibble(iris) %>%
  select( starts_with("Petal"), Species ) %>%
        filter( Petal.Length < 1.55 ) %>% 
         arrange(  Species )
```

## Q2

Long vs Wide data is a description on how a data table is created and stored. Long tables are built in a vertical sense, where each column indicates a different variable or parameter and each row represents 1 observation of data associated with some category of identifying variables and results. Wide tables are built in a horizontal sense, where each column instead indicates one observation of data and rows indicate the categorizing variables and resulting data information.

# Task 2

```{r}
# This code loads the glass.data file from URL with comma (,) Delimiter
glass_data <- 
  read_delim(
    "https://www4.stat.ncsu.edu/~online/datasets/glass.data",
    delim=",",
    col_names=c("Id", "RI", "Na", "Mg", "Al", "Si", "K", "Ca", "Ba", "Fe",
                "Glass_Type")
) %>% #create Glass type as factor text variable
        mutate(
          Glass_Type = factor( Glass_Type,
          levels = c(1, 2, 3, 5, 6, 7),
          labels = 
                   c("building_windows_float_processed",
                     "building_windows_non_float_processed",
                     "vehicle_windows_float_processed",
                     "containers", "tableware", "headlamps")
          ) 
          ) %>% #Filter data to Fe < 0.2 and select only types
                filter(
                  Fe < 0.2,
                  Glass_Type %in% c("tableware", "headlamps")
                  )

#print final dataset                         
glass_data                          

# This code reads in the yeast data from the URL
yeast_data <-  read_table(
  "https://www4.stat.ncsu.edu/~online/datasets/yeast.data", 
  col_names=c("seq_name", "mcg", "gvh", "alm", "mit", "erl", "pox",
              "vac", "nuc", "class")
) %>% #select to remove specified columns
    select( -seq_name, -nuc ) %>% #group for mean/median derivation
      group_by(class)  %>% #mutate to create specified columns using across()
        mutate(
          across(
            c(mcg, gvh, alm, mit, erl, pox, vac), 
            list(
            mean = ~mean(.x, na.rm = TRUE), 
            median = ~median(.x, na.rm = TRUE)
            )
            )
        )

#print final dataset                                
yeast_data     

# CREATE ALL_WINE DATASET AND BEGIN PIPING (copied from HW3)
  # This code reads the excel data for white Wine
  white_wine <- read_excel(
    "~/ST558/ST-558-Group-Assignments/Data/HW3/white-wine.xlsx",
    sheet=1
  )
  
  # this code creates a tibble of the excel sheet names.
  white_wine_names <- read_excel(
    "~/ST558/ST-558-Group-Assignments/Data/HW3/white-wine.xlsx",
    sheet=2
  )
  
  # assigning the column names of the white_wine dataset with the
  # string from white_wine_names
  colnames(white_wine) <- white_wine_names$Variables
  
  # assign White to the list of wines using mutate
  white_wine <- mutate(white_wine, type = "white")
  
  # This code reads in the red wine .CSV semicolon (;) delimited file,
  # assigns colnames, and assigns variable type a value of red
  red_wine <- read_delim(
    "https://www4.stat.ncsu.edu/~online/datasets/red-wine.csv",
    delim=";",
    col_names=c("fixed_acidity", "volatile_acidity", "citric_acid", 
                "residual_sugar", "chlorides", "free_sulfur_dioxide", 
                "total_sulfur_dioxide", "density", "pH", "sulphates", 
                "alcohol", "quality"),
    skip = 1
  ) %>% mutate(type = "red")

# Start Chaining/Piping of all_wine after merging wine datasets together
# combine white_wine and red_wine datasets using dply::bind_rows()
all_wine <- dplyr::bind_rows(list(red_wine, white_wine)) %>%
  #Filter data per specs.
  filter( quality > 6.5 & alcohol < 13.2) %>% 
    #sort data by desc quality
    arrange(desc(quality)) %>% 
      #select spec variables
      select(contains("acid"), alcohol, type, quality) %>%
        #group by quality
        group_by(quality) %>%
          mutate(
            across(
              c(alcohol), 
              list(
                mean = ~mean(.x, na.rm = TRUE), 
                median = ~median(.x, na.rm = TRUE)
              )
            )
          )
      
all_wine      
                              
```

# Task 3

```{r}
#Connect to database use DBI package
con <- dbConnect(RSQLite::SQLite(), "~/ST558/ST-558-Group-Assignments/Data/HW4/lahman.db")

#display tables
dbListTables(con)

#return all data from teams of 2015 year
teams2015 <- tbl(con,"Teams") %>% #select teams from year 2015
              filter( yearID == 2015 ) %>% #collect
               collect()

#display results
teams2015

#do the same with SQL
teams2015_v2 <- tbl(con,  
                    sql("select * from Teams where yearID = 2015")
                    ) %>%
                      collect()


#display results
teams2015_v2

#Return all HoF players.
hall_of_fame <- tbl(con,  
                    sql("select playerID, yearID, category Year from HallOfFame")
                    ) %>%
                      collect()


#display results
hall_of_fame

#join Hall of famer tables with players table.
#okay admitely I'm probably going a little beyond scope rn of what we've learned
#in class, but I use SQL with SAS for work a ton so I wanted to do some testing
#on it's capabilities. I'd be really interested in learning how to use SQL to 
#do this kind of handling on objects in the R environment!
Players_HoF <- tbl(con,  
                   sql("select a.nameFirst, a.nameLast, b.playerID, b.yearID, 
                        b.category 
                          from People as a
                            inner join HallOfFame as b
                              on a.playerID=b.playerID")
                    ) %>%
                      collect()

#display results
Players_HoF

#return manager ID 
manager_winLoss <- tbl(con, "Managers") |>
                    select(playerID, G, W, L) |>
                    group_by(playerID) |>
                    summarize(G_managed = sum(G, na.rm = TRUE),
                              Total_W = sum(W, na.rm = TRUE),
                              Total_L = sum(L, na.rm = TRUE)) |> 
                    collect() |>
                    mutate('Total_W/G_managed' = Total_W / G_managed) |>
                    arrange(desc(Total_W/G_managed))

manager_winLoss

#Return information for only the
#people that managed a team and were inducted into the hall of fame
manager_HoF <- inner_join(Players_HoF, manager_winLoss, by = "playerID")
manager_HoF
```
