---
title: "Riley_HW5"
format: html
---

Assignment Information\
ST 558\
Homework 5\
Individual\
Shannon Riley

```{r}
#| echo: false
#| output: false
# Load required packages
library(dplyr)    # to perform basic EDA
library(ggplot2)  # to make pretty plots, yay!
library(httr)     # to query an API
library(jsonlite) # to parse results of API query, tidyjson is another possible package to use
library(readr)    # to read in CSV data
library(stringr)  # to create clean column names
library(tidyr)    # to clean up tibbles
```

Task 1: Querying an API and a Tidy-Style Function
--

Use the *GET()* function from the *httr* package to return information about a topic in the news:
```{r}
queryInfo <- GET('https://newsapi.org/v2/everything?q=wildfires&from=2026-08-26&language=en&apiKey=ffb8439b27574e0190e19b8887522f24')
str(queryInfo, max.level = 1)
```

Parse the result of the API query and access the data frame that has the actual news article information:
```{r}
queryParse <- fromJSON(rawToChar(queryInfo$content))
wildfireInfo <- as_tibble(queryParse$articles)
wildfireInfo
```

```{r}
#| echo: false
#| output: false
# Unnest first column to improve readability of tibble:
# wildfireInfo |>
#   unnest(source, names_sep = '.')
```

Write a function that allows the user to query the API with the following character string inputs: title/subject search, time period, API key:\
**Inputs must be character strings with "%20" instead of spaces!**
```{r}
queryNewsAPI <- function(title_or_subject, date_to_search_from, API_key, ...) {
  if(!is.character(title_or_subject) | !is.character(date_to_search_from) | !is.character(API_key)){
    stop('Title or subject, date to search from, and API key must be character strings.')
  }
  queryInfo <- GET(paste('https://newsapi.org/v2/everything?q=', title_or_subject, '&from=', date_to_search_from, '&language=en&apiKey=', API_key, sep = ''))
  queryParse <- fromJSON(rawToChar(queryInfo$content))
  articleInfo <- as_tibble(queryParse$articles)
  return(articleInfo)
}
```

Test function twice:
```{r}
queryNewsAPI('squid', '08/26/2026', 'ffb8439b27574e0190e19b8887522f24')
```

```{r}
queryNewsAPI('Terry%20Pratchett', '08/26/2026', 'ffb8439b27574e0190e19b8887522f24')
```


Task 2: Summarizing Roller Coaster Data
--

Read in roller coaster data, clean up variable names, and convert variables to factors as needed:
```{r}
rc <- read_csv('https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv')
colnames(rc) <- str_to_camel(colnames(rc))
rc <- rc |>
  rename(numberInversions = ofInversions) |>
  mutate(censusRegionF = as.factor(censusRegion),
         ageGroupF = factor(ageGroup, levels = c('1:Older', '2:Recent', '3:Newest'), labels = c('Older', 'Recent', 'Newest')),
         scaleF = factor(scale, levels = c('Extreme', 'Thrill', 'Family')),
         typeF = as.factor(type),
         designF = as.factor(design),
         inversionsF = factor(inversions, levels = c('Yes', 'No')))
rc
```

View how the data are stored:
```{r}
str(rc)
```

Review basic details and statistics (if relevant) for each variable:
```{r}
summary(rc)
```

View just number of missing values per variable:
```{r}
colSums(is.na(rc))
```

Create contingency tables (one-way, two-way, three-way) for the factor variables:\
1) One-way
```{r}
table(rc$censusRegionF)
```
This contingency table is a one-way table showing how many observations (roller coasters) are in each region. For example, the West region has 122 roller coasters.
\
\
2) Two-way
```{r}
table(rc$typeF, rc$ageGroupF)
```
This second contingency table is a two-way table showing how many observations with a given type fall into each age group. For example, there are 35 wood roller coasters in the recent age group.
\
\
3) Three-way
```{r}
table(rc$designF, rc$scaleF, rc$inversionsF, useNA = 'always')
```
This contingency table is a three-way table (specifically a 3D array) showing how many roller coasters with or without inversions were a given design and scale. For example, there were 9 roller coasters with inversions, a flying design type and extreme scale.

Create a conditional two-way table using the *table()* function by:
\
1) Subsetting the data, then creating the table
```{r}
rcFilt <- rc |>
  filter(scaleF == 'Extreme')
table(rc$designF, rc$ageGroupF)
```
\
2) Creating a three-way table, then subsetting it
```{r}
table(rc$scaleF, rc$censusRegionF, rc$ageGroupF)
table(rc$scaleF, rc$censusRegionF, rc$ageGroupF)[, , 'Newest']
```

Create a two-way contingency table using *group_by()* and *summarize()* from *dplyr*, then use *pivot_wider()* to make it more readable:
```{r}
rc |>
  group_by(censusRegionF, designF) |>
  summarize(count = n()) |>
  pivot_wider(names_from = censusRegionF, values_from = count)
```

Create a stacked bar graph and a side-by-side bar graph:\
1) Stacked
```{r}
stackAgeRegion <- ggplot(rc, aes(x = ageGroupF, fill = censusRegionF)) +
  geom_bar(position = 'stack') +
  labs(x = 'Age Group', y = 'Count', fill = 'Region', title = 'Number of Roller Coasters by Age Group and Region')
stackAgeRegion
```
\
2) Side-by-side
```{r}
sideAgeRegion <- ggplot(rc, aes(x = ageGroupF, fill = censusRegionF)) +
  geom_bar(position = 'dodge') +
  labs(x = 'Age Group', y = 'Count', fill = 'Region', title = 'Number of Roller Coasters by Age Group and Region')
sideAgeRegion
```

