---
title: "Riley_HW5"
format: html
---

Assignment Information\
ST 558\
Homework 4\
Individual\
Shannon Riley

```{r}
#| echo: false
#| output: false
# Load required packages
library(dplyr)    # to perform basic EDA
library(httr)     # to query an API
library(jsonlite) # to parse results of API query, tidyjson is another possible package to use
library(readr)    # to read in CSV data
library(stringr)  # to create clean column names
library(tidyr)    # to clean up tibbles
```

Task 1: Querying an API and a Tidy-Style Function
--

Use the *GET()* function from the *httr* package to return information about a topic in the news:
```{r}
queryInfo <- GET('https://newsapi.org/v2/everything?q=wildfires&from=2026-08-20&language=en&apiKey=ffb8439b27574e0190e19b8887522f24')
str(queryInfo, max.level = 1)
```

Parse the result of the API query and access the data frame that has the actual news article information:
```{r}
queryParse <- fromJSON(rawToChar(queryInfo$content))
wildfireInfo <- as_tibble(queryParse$articles)
wildfireInfo
```

```{r}
#| echo: false
#| output: false
# Unnest first column to improve readability of tibble:
# wildfireInfo |>
#   unnest(source, names_sep = '.')
```

Write a function that allows the user to query the API with the following character string inputs: title/subject search, time period, API key:
**Inputs must be character strings with %20 instead of spaces!*
```{r}
queryNewsAPI <- function(title_or_subject, date_to_search_from, API_key, ...) {
  if(!is.character(title_or_subject) | !is.character(date_to_search_from) | !is.character(API_key)){
    stop('Title or subject, date to search from, and API key must be character strings.')
  }
  queryInfo <- GET(paste('https://newsapi.org/v2/everything?q=', title_or_subject, '&from=', date_to_search_from, '&language=en&apiKey=', API_key, sep = ''))
  queryParse <- fromJSON(rawToChar(queryInfo$content))
  articleInfo <- as_tibble(queryParse$articles)
  return(articleInfo)
}
```

Didn't work - delete before submitting
```{r}
queryNewsAPI <- function(title_or_subject, date_to_search_from, API_key, ...) {
  result <- tryCatch({
    if (!is.character(title_or_subject) | !is.character(date_to_search_from) | !is.character(API_key)){
    stop('Title or subject, date to search from, and API key must be character strings.')
    }
    queryInfo <- GET(paste('https://newsapi.org/v2/everything?q=', title_or_subject, '&from=', date_to_search_from, '&language=en&apiKey=', API_key, sep = ''))
  queryParse <- fromJSON(rawToChar(queryInfo$content))
  articleInfo <- as_tibble(queryParse$articles)
  }, error = function(e) {
    message("Error: ", e$message)
    NA  # Returning NA in case of an error
  })
  return(result)
}
```

Test function twice:
```{r}
queryNewsAPI('squid', '08/20/2026', 'ffb8439b27574e0190e19b8887522f24')
```

```{r}
queryNewsAPI('Terry%20Pratchett', '08/20/2026', 'ffb8439b27574e0190e19b8887522f24')
```


Task 2: Summarizing Roller Coaster Data
--

Read in roller coaster data, clean up variable names, and convert variables to factors as needed
```{r}
rc <- read_csv('https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv')
colnames(rc) <- str_to_camel(colnames(rc))
rc |>
  rename(numberInversions = ofInversions)
```

