---
title: "Querying APIs and EDA"
author: "Cody Ashby"
format: html
editor: visual
---

## Task One: Querying an API and a Tidy-Style Function

1.  The first part of this assignment will focus on connecting to an API. We'll need to upload the `httr` package as well as attain an API key. The topic of interest is one that's been in the news not just the last few weeks, but the last few years! Since it's a really hot topic, we'll just look at articles posted after Labor Day.

```{r}
library(httr)

#Use this as the API key: "5970f6f1267047469fb8667ecde70a24"

test_url=GET("https://newsapi.org/v2/everything?q=inflation&from=2026-09-08&language=en&pageSize=50&apiKey=5970f6f1267047469fb8667ecde70a24")
test_url
```

Below is the structure of this url.

```{r}
str(test_url,max.level = 1)
```

2.  Sticking to what's been demonstrated in lectures, we'll upload the `jsonlite` package.

```{r}
library(jsonlite)
```

We'll need to use parsing methods here and convert any raw information from this `json`-type object to a character format and further convert it to a tibble.

```{r}
parsed_url=fromJSON(rawToChar(test_url$content))

articles=as_tibble(parsed_url$articles)
articles
```

Wait! The first column shouldn't be a collection of data frames! Let's remove that column.

```{r}
simplified_articles=articles[-1]
simplified_articles
```

There! That's better.

3.  Next, we'll create a function that allows a user to input a topic, a date, and an API key.

```{r}
search_it <- function(topic,date,APIkey){
  if(!is.character(topic)|!is.character(date)|!is.character(APIkey)){
    stop("At least one of these is NOT a string. Use quotation marks and '%20' instead of a space for better results. Also, make sure the date is of the format yyyy-mm-dd.")
  }
  lengthy_url = GET(paste("https://newsapi.org/v2/everything?q=",
                          topic,"&from=",date,"&language=en&pageSize=50&apiKey=",APIkey,sep=''))
  parsed_search = fromJSON(rawToChar(lengthy_url$content))
  top_articles = as.tibble(parsed_search$articles)
  return(top_articles[-1])
}
```

Let's try this function out!

```{r}
search_it('groceries','2026-09-18','5970f6f1267047469fb8667ecde70a24')
```

```{r}
search_it('midterm%20elections','2026-09-18','5970f6f1267047469fb8667ecde70a24')
```

## Task Two: Summarizing Roller Coaster Data

Before diving in, let's upload the `tidyverse` library and read in this data set.

```{r}
library(tidyverse)
roller_coaster_data=read_csv("https://www4.stat.ncsu.edu/online/datasets/635USrollercoasters.csv")

roller_coaster_data
```

There's also a good handful of categorical variables; let's use `factor` methods.

```{r}
roller_coaster_data |>
  mutate(Region=as.factor(`Census Region`),
         AgeGroup=as.factor(AgeGroup),
         Scale=as.factor(Scale),
         Type=as.factor(Type),
         Design=as.factor(Design),
         `Inversions?`=as.factor(`Inversions?`))
```

#### Basic EDA

4.  Let's look at the structure of this object to see how it's stored and if there's anything that makes no sense.

```{r}
str(roller_coaster_data)
```

5.  The `colSums` function will now be used along with the `is.na` function to determine which variables have missing values, if any.

```{r}
colSums(is.na(roller_coaster_data))
```

#### Categorical Variables

6.  Let's make a few contingency tables based on up to three categories: the age, type and region of each roller coaster.

```{r}
# One-way table for age of coasters
table(roller_coaster_data$AgeGroup)
# Two-way table for age of coasters and type
table(roller_coaster_data$AgeGroup,roller_coaster_data$Type)
# Three-way table for age of coasters, type, and census region
table(roller_coaster_data$AgeGroup,roller_coaster_data$Type,roller_coaster_data$`Census Region`)
```

Some extra commentary for these contingency tables will go here.

7.  Next, we'll use two methods for constructing a conditional two-way table. One will involve subsetting the data before creating the table, and the other will use a subset of a three-way table.

```{r}
# Focus on type and scale of coasters in South region
southern_coasters <- roller_coaster_data |>
  filter(`Census Region`=="South")

# Two-way table from subsetted data
table(southern_coasters$Type,southern_coasters$Scale)

# Subset of three-way table
table(roller_coaster_data$Type,roller_coaster_data$Scale,roller_coaster_data$`Census Region`)[,,"South"]
```

8.  We can also use `dplyr` to create a two-way contingency table; however, we'll need to use a pivoting method so that it actually looks like a table.

```{r}
roller_coaster_data |>
  group_by(`Census Region`,Type) |>
  summarize(count=n()) |>
  pivot_wider(names_from = `Census Region`,values_from = count)
```

9.  Lastly, we'll create two types of bar graphs, one stacked and one side-by-side

```{r}
ggplot(roller_coaster_data,aes(x=`Census Region`,fill=Type)) +
  geom_bar() + 
  labs(x="Region",y="Number",fill="Region",title="Number of Roller Coasters by Type and Region")

ggplot(roller_coaster_data,aes(x=`Census Region`,fill=Type)) +
  geom_bar(position="dodge") + 
  labs(x="Region",y="Number",fill="Region",title="Number of Roller Coasters by Type and Region")
```
