---
title: "Xiao_HW5"
format: html
editor: visual
---

# Homework 5

#### Task 1 - Querying an API and a Tidy-Style Function

```{r}
#Install packages as necessary
#install.packages("httr")
#install.packages("jsonlite")
```

```{r}
#Load Packages and API Key
library(httr)
library(jsonlite)
library(tidyverse)

news_api_key <- "74a38983f4844f5a8891b351f068b813"
```

1\. Use GET() from the httr package to return information about a topic that you are interested in that has been in the news lately (store the result as an R object). Note: We can only look 30 days into the past with a free account.

```{r}
#Query the API using 

news_response <- GET(
  url = "https://newsapi.org/v2/everything",
  query = list(
    q = "artificial intelligence",
    from = Sys.Date() - 14, #14 Days Ago
    lanaguage = "en",
    pageSize = 50,
    apiKey = news_api_key
  )
)

news_response
```

2\. Parse what is returned and find your way to the data frame that has the actual article information in it (check content). Note the first column should be a list column!
```{r}
#Extract news content by parsing the response into an R object
news_content <- content(news_response, as = "parsed", simplifyVector = TRUE)

#Extract articles from the above content
news_articles <- news_content$articles

#Structure of resulting article data
str(news_articles)

#Display first few articles
head(news_articles)

#The resulting object contains info about the articles returned by the query, including the source, author, title, description, URL, publication date, and available article content

#Note that the source variable is stored as a list column because each source contains multiple pieces of information
```

3\. Now write a quick function that allows the user to easily query this API. The inputs to the function should be the title/subject to search for (string), a time period to search from (string - you’ll search from that time until the present), and an API key.

```{r}

```


###### Use your function twice to show it works!

#### Task 2 - Summarizing Roller Coaster Data

4\. Look at how the data is stored and see if everything makes sense.

5\. Document any missing values in the data.

#### Categorical Variables

6\. Create a one-way contingency table, a two-way contingency table, and a three-way contingency table for some of the factor variables you created previously. Use table() to accomplish this.

-Interpret a number from each resulting table (that is, pick out a value produced and explain what that value means.)

7\. Create a conditional two-way table using table(). That is, condition on one variable’s setting and
create a two-way table. Do this using two different methods:
- Once, by subsetting the data (say with filter()) and then creating the two-way table
- Once, by creating a three-way table and subsetting it

8\. Create a two-way contingency table using group_by() and summarize() from dplyr. Then use pivot_wider() to make the result more like the output from table().

9\. Create a stacked bar graph and a side-by-side bar graph. Give relevant x and y labels, and a title for the plots.